\chapter{Package \texttt{clremoval}}


This package provides routines for finding line harmonics,
generating a reference interference signal, and removing
all the interference harmonics. As an example, these routines can be usefull
for removing all those lines corresponding to the electricity supply
 frequency and its harmonics.

The technique used  is the so-called coherent line removal
({\sc clr}) \cite{Sintes:1998}-\cite{Sintes:2000}.
 {\sc clr} is an algorithm able to remove interference present in the
 data while preserving the stochastic detector noise.  {\sc clr}
 works when the interference is present in many harmonics, as long as
   they remain coherent with one another.
{\sc clr} can remove the external
 interference without removing any \lq single line' signal buried by the
 harmonics. The algorithm works even when the interference frequency changes.
{\sc clr} can be used to remove all harmonics of periodic or
broad-band signals (e.g., those which change frequency in time), even when
there is no external reference source.  {\sc clr} requires little
 a priori knowledge of the signals we want to remove.

The package  is organized under the header \verb@CLR.h@ and the modules
  \verb@HarmonicFinder.c@,  \verb@RefInterference.c@ and
 \verb@CleanAll.c@.



\newpage\input{CLRH}


\newpage\begin{thebibliography}{0}

\bibitem{Sintes:1998}
  A.~M. Sintes, B.~F. Schutz, Phys. Rev. D\textbf{58}, 122003 (1998)

\bibitem{Sintes:1998b}
  A.~M. Sintes, B.~F. Schutz,
 \textit{Removal of interference from gravitational wave spectrum}, in
 \textit{Proceedings of the 2nd workshop on Gravitational Wave
  Data Analysis}.
  Edts. M. Davier and P. Hello. Editions Fronti\`eres,
  France (1998) pp. 255-260

\bibitem{Sintes:1998c}
  A.~M. Sintes, B.~F. Schutz,
  \textit{Removal of interference from external coherent signals}, in
  \textit{Laser Interferometer Space Antenna:
  Second international LISA symposium on gravitational waves}.
  Edt. W. Folkner. American Institute of Physics, New York (1998) pp. 135-140

\bibitem{Sintes:1999}
  A.~M. Sintes, B.~F. Schutz, Phys. Rev. D\textbf{60}, 062001 (1999)

\bibitem{Sintes:2000}
  A.~M. Sintes, B.~F. Schutz,
  \textit{Removing Line Interference from Gravitational Wave
   Interferometer Data}, in
  \textit{Gravitational Wave Detection II}.
  Edts. S. Kawamura and N. Mio. Frontiers Science Series n. 32;
  Universal Academy Press, Inc, Tokyo  (2000) pp. 273-278

\end{thebibliography}

