\section{Program \texttt{lalapps\_hello}}
\label{program:lalapps-hello}
\idx[Program]{lalapps\_hello}

\begin{entry}

\item[Name]
\verb$lal_hello$ --- prints ``hello LSC!''

\item[Synopsis]
\verb$lal_hello$ [\verb$-h$] [\verb$-V$] [\verb$-v$]
[\verb$-d$ \textit{dbglvl}] [\verb$-o$ \textit{outfile}]

\item[Description]
\verb$lal_hello$ prints ``hello LSC!'' to the screen or to an output file.

\item[Options]\leavevmode
\begin{entry}
\item[\texttt{-h}]
Print a help message.
\item[\texttt{-V}]
Print the version information.
\item[\texttt{-v}]
Verbose output.
\item[\texttt{-d} \textit{dbglvl}]
Set LAL debug level to \textit{dbglvl}.
\item[\texttt{-o} \textit{outfile}]
Write the output to file \textit{outfile}.
\end{entry}

\item[Debug levels]
The LAL debug level can be specified as an integer or as a string of flags:
\begin{entry}
\item[\texttt{NDEBUG}]
No debugging information is printed and memory debugging code is disabled.
\item[\texttt{ERROR}]
Error messages are printed.
\item[\texttt{WARNING}]
Warning messages are printed.
\item[\texttt{INFO}]
Information messages are printed.
\item[\texttt{TRACE}]
Function call tracing messages are printed.
\item[\texttt{MEMINFO}]
Memory  allocation  information messages are printed.
\item[\texttt{MEMDBG}]
Debugging of memory allocation routines is enabled but no messages are printed.
\end{entry}
The following composite levels are available:
\begin{entry}
\item[\texttt{MSGLVL1}]
Equivalent to \verb$ERROR$
\item[\texttt{MSGLVL2}]
Equivalent to \verb$ERROR | WARNING$
\item[\texttt{MSGLVL3}]
Equivalent to \verb$ERROR | WARNING | INFO$
\item[\texttt{ALLDBG}]
All debugging messages are printed.
\end{entry}

For example, the command
\begin{indented}
\verb$lal_hello -d "ERROR | INFO"$
\end{indented}
will set the debug level so that error and information messages are printed.

\item[Environment]\leavevmode

\begin{entry}
\item[\texttt{LAL\_DEBUG\_LEVEL}]
Default LAL debug level to use.
\end{entry}

\item[Author]
Jolien Creighton

\end{entry}
