\chapter{Inspiral Search Programs}
\label{chapter:inspiral}

This section of \textsc{LALApps} contains programs that can be used to search
interferometer data for inspiral signals using templated matched filtering and
associated veto strategies.

\section{Description of the Components of the Pipeline}

A short description of each possible part of the inspiral pipeline follows. 
The more detailed description of each command-line argument can be found in 
the specific pages for each part of the code.

\subsection{Template Bank Generation}

\texttt{lalapps\_tmpltbank}: Calculates the power spectrum and generates a 
template bank for an analysis
chunk of LIGO or GEO data. The bank can be for any of the different kinds of
inspiral signals: TaylorF2, BCV etc. The output is written in an aml file.

There is also the capability of outputting the raw data,
the response function and the uncalibrated power spectrum in
frame files.

\subsection{Matched Filtering and Injections}

\texttt{lalapps\_inspiral}: Estimates the power spectrum and performs matched 
filtering for inspiral signals on LIGO or GEO data. The filter can be for
either binary neutron star inspirals or for spinning or non-spinning black 
hole inspirals.
It also has the capability of doing software injections of 
inspiral signals on the data. The resulting triggers are written in an xml
file.

There is also the capability of outputting the raw data before and after the
injections, the response function, the uncalibrated power spectrum, the snr,
the chisq time series, and the complex C time series required by the coherent
code, all in frame files.

\texttt{lalapps\_coherent\_inspiral}: Takes C data frame files as 
outputted by inspiral.c for up to 4 detectors and constructs the optimal 
coherent SNR time series.  Triggers are written to an xml file and the time 
series is written out as a frame file.

\subsection{Triggered Template Bank, Trigger Coincidence and Time Slides}

\texttt{lalapps\_inca}: Currently performs three different tasks:

Triggered template bank generation: Given the triggers from one interferometer
in a \verb$sngl_inspiral$ table, it generates a triggered tempalte bank to be
used in filtering the data from a second interferometer. 

Trigger coincidence: Given the triggers from two interferometers, it checks
which of those are time-coincident and coincident in mass (for the BNS search)
or in the parameters psi0 and psi3 (for the BCV search).

Time slides: Time slides the data by a given time and performs coincidence,
as described above.

All results are written in xml files.

\texttt{lalapps\_thinca}: This code is intended as a generalization of inca.  It
is not fully implemented at the present time.  For now, it can take in triggers
from any number of instruments and return triggers which are coincident in two
of the instruments.  The coincidence test is performed over time and mass
parameters.  Furthermore, the events are labelled with a unique id in order to
facilitate post processing.  

All results are written in xml files.

\texttt{lalapps\_trigbank}:  This code generates a triggered template bank.
Given an input of a \verb$sngl_inspiral$ table, it generates a triggered
tempalte bank to be used in filtering the data from a second interferometer. 
All results are written in xml files.

\subsection{Trigger Summary and Injection Analysis}

\texttt{lalapps\_sire}: Currently performs two different tasks:

Summary and clustering of inspiral triggers: Using the xml files with the
inspiral triggers, it summarizes them and time-clusters them, using a
specified clustering algorithm.

Injection analysis: Performs time-coincidence between  the inspiral triggers 
and a list of injection parameters. An injection is considered "found" if
there is an inspiral trigger that is time-coincident with it, within a 
specified time interval.

All results are written in xml files.

\subsection{Injection Parameter Generation}

\texttt{lalapps\_inspinj}: Given a mass-file, it generates the mass,
  distance and time parameters of BNS signals to be injected in the data.
  It currently generates the signals coming from locations within the
  galaxies of the Local Group, but the ability to do different distributions
  will be added soon.

  The result is an xml file.

\texttt{lalapps\_bbhinj}: It generates the mass,
  distance and time parameters of inspiral signals to be injected in the data.
  Despite its name, it can be used for BNS and for BBH parameter generation.

  The result is an xml file.

\subsection{Injection in Frames}

\texttt{lalapps\_inspfrinj}: Given a frame file with LIGO or GEO data, it
injects inspiral signals on the data.


\subsection{Splitting Large Template Banks}

\texttt{lalapps\_splitbank}: Given a template bank in an xml file, it splits
that bank into smaller banks.

\clearpage
\input{inspiral_pipe}

\clearpage
\input{inspiral_hipe}

\clearpage
\input{tmpltbank}

\clearpage
\input{inspiral}

\clearpage
\input{coherent_inspiral}

\clearpage
\input{inca}

\clearpage
\input{thinca}

\clearpage
\input{trigbank}

\clearpage
\input{sire}

\clearpage
\input{inspinj}

\clearpage
\input{bbhinj}

\clearpage
\input{inspfrinj}

\clearpage
\input{inspmultiawg}

\clearpage
\input{splitbank}

%\clearpage
%\input{BankEfficiency.tex}
%\clearpage
%\input{BankNumber.tex}
