\chapter{Package \texttt{tdfilter}}

This package covers LAL routines for constructing and applying digital
time-domain filters.  It is organized under three headers.  The
\verb@ZPGFilter.h@ header provides routines for manipulating filters
in the ``zeros, poles, gain'' representation, which is typically the
simplest way of representing a filter response.  These routines create
and destroy ZPG filters, and can transform the complex variable used
to represent them.  The \verb@IIRFilter.h@ header provides routines
for creating actual time-domain filters from the ZPG representation,
and applying these filters to data.  The \verb@BandPassTimeSeries.h@
header provides routines an actual implementation of these utilities
to the specific task of high- or low-pass filtering of a data stream.
These routines also serve as an example for the more general task of
designing time-domain filters with any desired frequency response.

The module \verb@ButterworthTimeSeries.c@ provides specific advice and
guidelines for building a numerically stable time-domain filter, but
the general procedure is as follows.  (1) Decide on the desired filter
response, and express it as a rational function of the frequency
variable $w=\tan(\pi f\Delta t)$ (which maps the Nyquist interval onto
the positive real axis).  (2) Factorize this rational function into
zeros and poles, restricting oneself to the upper half of the $w$
complex plane.  Assign these to one or more objects of type
\verb@<datatype>ZPGFilter@.  (3) Use \verb@WToZ<datatype>ZPGFilter()@
to transform the filter to the more conventional $z=\exp(2\pi if\Delta
t)$ frequency variable.  (4) Use the routines in \verb@IIRFilter.h@ to
create IIR filters from the ZPG filters, and to apply them to data.

\newpage\input{ZPGFilterH}
\newpage\input{IIRFilterH}
\newpage\input{BandPassTimeSeriesH}

\newpage\input{LPCH}
\newpage\input{LPCC}
