\begin{manpage}{\prog{ligolw\_binjfind}}

\manname{\prog{ligolw\_binjfind}}{A program for finding burst injections.}

\mansynopsis{\prog{ligolw\_binjfind} \{\progarg{-c},\progarg{--match-algorithm}\} \progparm{algorithm} [options] [file [file ...]]}

\begin{mansection}{Description}
\prog{ligolw\_binjfind} applies a burst injection identification algorithm
to one or more LIGO Light Weight XML files, each containing a list of burst
triggers and a list of software injections.  If no file names are provided
on the command line, then input is read from \file{stdin}, and output
written to \file{stdout}.  Any file whose name ends in ``\file{.gz}'' is
assumed to be gzip-compressed.

The results of the search are recorded using the standard coincidence table
infrastructure.
\end{mansection}

\begin{mansection}{Options}
\begin{description}
\item[\progarg{--comment} \progparm{text}] Set the string to be recorded as
the comment in the job metadata added to the files.  The default is ``''.

\item[\progarg{-f},\progarg{--force}] Normally input files whose internal
metadata indicates they have already been processed will be skipped.  This
behaviour facilitates the use of \prog{ligolw\_binjfind} in Condor DAGs,
where it is convenient to have one job process many files but where a job
may be evicted and re-started, causing it to re-read each file.  Setting
\progarg{--force} forces files to be re-processed.

\item[\progarg{-h},\progarg{--help}] Show a help message and exit.

\item[\progarg{-c},\progarg{--match-algorithm} \progparm{algorithm}] Set
the algorithm to be used in comparing burst triggers to injections.  There
are currently two algorithms defined, \progarg{excesspower} and
\progarg{stringcusp}.  The former is suitable for use in comparing the
triggers generated by \prog{lalapps\_power} to a list of burst injections,
while the latter is for use in comparing the triggers generated by
\prog{lalapps\_StringSearch}.

\item[\progarg{--verbose}] Be verbose.

\item[\progarg{--version}] Show program's version number and exit.

\end{description}
\end{mansection}

\begin{mansection}{See Also}
\prog{ligolw\_add}, \prog{lalapps\_binj}, \prog{ligolw\_burca}
\end{mansection}

\end{manpage}
